The Noisy Work of Uncertainty Visualisation Research: A Review

### 2024-08-19 paper.qmd
spent ages making a GT table and it doesnt even work in the PDF file. I hate this.

# load Libraries
library(tidyverse)
library(biscale)
library(RColorBrewer)
library(scales)
library(sf)
library(ggrepel)
library(urbnmapr)
```

@@

## Example Data
Throughout this review, we will be looking at

@@

my_map_data <- my_map_data |>
  dplyr::mutate(temp = 28 - ((scale(cent_long))^(2) + (scale(cent_lat))^(2))[,1], # trend
         notrend = rnorm(n), # no trend
         highvar = runif(n, min=2, max=4), # high variance
         lowvar = runif(n, min=0, max=2), # low variance
         ) |>
  pivot_longer(cols=highvar:lowvar, names_to = "variance_class", values_to = "variance") |>
  # add bivariate classes to data
  mutate(bitemp = cut(temp, breaks=breaks, labels=seq(8)),
         bivar = cut(variance, breaks=0:4, labels=seq(4)),
         biclass = paste(bitemp, bivar, sep="-")
  )

@@

- add map that is just outline
- include function used to design temperature

Due to the fields origins and focus in geospatial information visualisation, there have been a large number of suggested variations on the choropleth map that allow authors to include uncertainty. We will use these maps to provide simple examples for each approach that can be easily compared to the "no uncertainty" choropleth map to better understand the costs and benefits of each approach. Despite the example of each method focusing on variations of the choropleth map, it is important to understand that the approaches we are discussing are universal and are not unique to maps.

>>>>

# load Libraries
library(tidyverse)
library(biscale)
library(RColorBrewer)
library(scales)
library(sf)
library(ggrepel)
library(urbnmapr)
library(gt)
```

@@

## Spatial examples
There is an endless number of uncertainty visualisations and it would be impossible to show all of them in this review. While the theoretical approach we outline is useful to all uncertainty visualisations, regardless of their application, we feel that showing the variations on a single visualisation will help isolate the ideas we are trying to convey. Due to the fields origins and focus in geospatial information visualisation, there have been a large number of suggested variations on the choropleth map that allow authors to include uncertainty. We will use these maps to highlight the costs and benefits of each approach.

@fig-data shows the first six rows of our example data, along with the geographical boundaries of our data set.  The temperature variable was generated using a quadratic spatial trend, that is $Temperature = 28-SLongitude^2 + SLatitude^2$ where the $SLongitide$ and $SLatitude$ are the scaled (to standard normal) longitude and latitude of the county's centroid. The high and low cases for the sampling variance are randomly sampled from a uniform distribution. The low standard errors are drawn from a $U_{[0,1]}$ distribution, while the high standard error cases are drawn from a a $U_{[1,2]}$ distribution. This is the data we will be using in our spatial uncertaity examples for the rest of the paper.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-data
#| layout-nrow: 1
#| fig-cap: "The"
#| cap-location: "bottom"

@@

my_map_data <- my_map_data |>
  dplyr::mutate(temp = 28 - ((scale(cent_long))^(2) + (scale(cent_lat))^(2))[,1], # trend
         notrend = rnorm(n), # no trend
         highvar = runif(n, min=1, max=2), # high variance
         lowvar = runif(n, min=0, max=1), # low variance
         ) |>
  mutate(highvar = highvar^2,
         lowvar = lowvar^2,
         count_id = row_number()) |>
  pivot_longer(cols=highvar:lowvar, names_to = "variance_class", values_to = "variance") |>
  # add bivariate classes to data
  mutate(bitemp = cut(temp, breaks=breaks, labels=seq(8)),
         bivar = cut(variance, breaks=0:4, labels=seq(4)),
         biclass = paste(bitemp, bivar, sep="-"))

# Make nice table
example_table <- my_map_data |>
  mutate(variance = sqrt(variance)) |>
  select(c(count_id, county_name, temp, variance_class, variance)) |>
  as_tibble()|>
  pivot_wider(id_cols=c(count_id, county_name, temp,), 
              names_from = variance_class, 
              values_from = variance) |>
  head()|>
  gt() |>
   tab_header(
    title = "Average Daily High Temperatures of Iowa Counties",
    subtitle = "Psudo-data with high standard error and low standard error cases") |>
  tab_spanner(
    label = "Standard Error",
    columns = c(highvar, lowvar)
  ) |>
  cols_label( 
    count_id = "ID", 
    county_name  = "County",
    temp = "Temperature (°C)",
    highvar = "High",
    lowvar = "Low"
    ) |>
  fmt_number(columns = c(temp, highvar, lowvar)) |>
  data_color(
    columns = temp,
    method = "numeric",
    palette = "Oranges",
    domain = c(21, 29)
  ) |>
  data_color(
    columns = c(highvar, lowvar),
    method = "numeric",
    palette = "Greens",
    domain = c(0, 3)
  ) |>
  as_latex()

# make blank map 
example_map <- my_map_data |>
  filter(variance_class=="lowvar") |>
  ggplot() +
  geom_sf(aes(geometry = geometry), fill="white") + 
  #geom_text(aes(x=cent_long, y=cent_lat, label=count_id), size=2) +
  theme_void()

example_table

example_map

@@
